---
title: "Applied Predictive Modeling"
subtitle: "Chapter 5: Measuring Performance in Regression Models"
author: "Santiago Toso"
output: 
  html_document:
  number_sections: yes
  theme: cosmo
  toc: yes
  toc_float: yes
---

The following sections will reference functions from the `caret` package.

To compute model performance, the observed and predicted outcomes should be stored in vectors. For regression, the vectors should be numeric. Here two vectors are manually created to illustrate the techniques:

```{r}
observed <- c(.22, .83, -.12, .89, -.23, -1.3, -.15, -1.4, .62, .99, -.18, .32, .34, -.30, .04, -.87, .55, -1.3, -1.15, .2)

predicted <- c(.24, .78, -.66, .53, .7, -.75, -.41, -.43, .49, .79, -1.19, .06, .75, -.07, .43, -.42, -.25, -.64, -1.26, -.07)

residualValues <- observed - predicted
summary(residualValues)
```

An important step in evaluating the quality of the model is to visualize the results. First, a plot of the observed values against the predicted values helps one to understand how well the model fits. Also, a plot of the residuals versus the predicted values can help uncover systematic patterns in the model predictions.

```{r}
axisRange <- extendrange(c(observed, predicted))
plot(observed, predicted,
     ylim = axisRange,
     xlim = axisRange)

#Add 45 degrees reference line
abline(0, 1, col = "darkgrey", lty = 2)

#Predicted values versus residuals
plot(predicted, residualValues, ylab = "residual")
abline(0, 1, col = "darkgrey", lty = 2)
```

The `caret` package contains functions for calculating the RMSE and R^2 value:

```{r}
library(caret)
R2(predicted, observed)
RMSE(predicted, observed)
```

There are different formulas for R^2. By default, `R2` uses the square of the correlation coefficient. Base `R` contains a function to compute the correlation, including Spearman's rank correlation.

```{r}
# Simple correlation
cor(predicted, observed)
cor(predicted, observed)^2

# Rank correlation
cor(predicted, observed, method = 'spearman')
```

